### 2021-07-07 doc/thesis/chapters/dsp.tex
downsampling

The main advantage of direct-conversion receivers is, that their output is an image-free signal.
This is due to the properties of the mixer.
The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer output, and the flicker noise, also called $\frac{1}{f}$-noise, which induces strong noise at low frequencies \cite{PassosFábio2020AHSo}.
These effects are not explained into more detail here, since this is not the focus of this thesis.

@@

In words: The sampling frequency must be higher than twice the frequency of the signals' highest frequency content.
Any digital signal processing technique must hold the specification of the sampling theorem \cite{AlessioSilviaMaria2016DSPa}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Oversampling / Sample Rate Reduction}
%TODO: explain DSP functions that are actually used in Matlab FM demod
Example: \\
In case the multiplex signal needs to be stored in a file with 44.1 kHz samplerate, in order to avoid aliasing, a lowpass with a cutoff frequency of 15 kHz and sufficient attenutation before 19 kHz needs to be inserted before the decimation and file output.

>>>>

The main advantage of direct-conversion receivers is, that their output is an image-free signal.
This is due to the properties of the mixer.
The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer output, and the flicker noise, also called \textit{1/f}-noise, which induces strong noise at low frequencies \cite{PassosFábio2020AHSo}.
These effects are not explained into more detail here, since this is not the focus of this thesis.

@@

In words: The sampling frequency must be higher than twice the frequency of the signals' highest frequency content.
Any digital signal processing technique must hold the specification of the sampling theorem \cite[chpt. 4.2]{AlessioSilviaMaria2016DSPa}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sample Rate Reduction}
%TODO: explain DSP functions that are actually used in Matlab FM demod

Sample rate reduction is also called downsampling.
Downsampling is a way to reduce the amount of data that needs to be processed, without losing any of the signals' information.
In some applications, the digital signal is oversampled by a large factor.
However, if the frequency spectrum of the desired signal is a narrow band, a high oversampling rate is not necessary and the signal can be downsampled without losing any information content.
To guarantee this, the Nyquist sampling theorem must still be met after the downsampling process.

A lower sample rate is especially useful for efficient filter design.
Factors such as the transition bandwidth or the width of the passband are important there.
Short transition bandwidths, or a narrow passband require a high order of the filter to fulfill the specifications, which translates into a high computational effort.
Thus, it is of advantage to downsample the input, in order to generate better conditions for the filter design.
Downsampling spreads the spectrum of interest, in terms of the relative bandwidth between passband and sampling frequency.
Therefore, the filter can be designed with wider transition bandwidth and passband, which significantly reduces the required filter order.

As an example, assuming the signal of interest has a bandwidth of 20 kHz.
The Nyquist theorem thus requires a remaining sampling rate of at least twice the signals' frequency, so 40 kHz.
The signal is sampled with an oversampling factor of 10 - meaning a sample rate of 200 kHz.
A downsampling factor of 4 is now applied, which results in a sample rate of 50 kHz.
The resulting signal is now downsampled and can be processed at this lower data rate, while none of the original data content is lost.\\

The process consists of two parts: filtering and skipping samples.
Let us assume downsampling by a factor of $M=4$.
In the first step, a low-pass filter with a normalized cut-off frequency of \textit{1/M} is applied.
This is required in order to avoid aliasing later in the process.
The second step simply skips a number of M-1 samples in the sequence.

More detailed information can be found in \cite[chpt. 6.9]{AlessioSilviaMaria2016DSPa}, \cite[chpt. 4.1]{GopiE.S2018MDSP}
\cite[chpt. 10.2.2]{ThyagarajanK.S2019ItDS}.
